---
title: "Untitled"
author: "cathy"
date: "2024-12-06"
output: html_document
---
```{r}
library(tidyverse)
library(janitor)
library(knitr)
library(gridExtra)
library(reshape2)
library(plotly)
```
```


```{r}
commuting_df = read.csv("data/commuting.csv", na = c("NA",".","")) |> 
  janitor::clean_names() |> 
  rename(geo_type = geo_type_desc) |> 
  filter(
    geo_type %in% c("UHF 42", "Citywide", "Borough"))
```

draw graphs for commuting percent
first for counts graph
```{r}
uhf42_data = commuting_df  |> 
  filter(geo_type == "UHF 42")  |> 
  select(geography, bicycle_number, car_truck_or_van_number, 
         public_transportation_number, walked_number,
         bicycle_percent, car_truck_or_van_percent,
         public_transportation_percent, walked_percent)
         
uhf42_counts <- uhf42_data |> 
  select(geography, bicycle_number, car_truck_or_van_number, 
         public_transportation_number, walked_number)  |> 
  melt(id.vars = "geography", 
       variable.name = "Commute_Mode", 
       value.name = "Count")
       
  uhf42_counts$Count = as.numeric(gsub(",", "", uhf42_counts$Count))    

custom_colors = c("lightpink", "darkolivegreen", "gold", "thistle")


plot_counts <- plot_ly(uhf42_counts, 
                       x = ~geography, y = ~Count, 
                       color = ~Commute_Mode,  type = "bar",
                       colors = custom_colors) %>%
  layout(title = list(text = "Commuting Mode Counts by UHF 42 Neighborhoods", font = list(size = 9)),
         xaxis = list(title = "Neighborhood", tickangle = 45),
         yaxis = list(title = "Count"),
         barmode = "stack",
         legend = list(title = list(text = "Commute Mode"),
                       font = list(size = 7)),
         font = list(size = 7))
         
plot_counts
```

second for percentage graph
```{r}
uhf42_percentages = uhf42_data  |> 
  select(geography, bicycle_percent, car_truck_or_van_percent,
         public_transportation_percent, walked_percent)  |> 
  melt(id.vars = "geography", variable.name = "Commute_Mode", value.name = "Percent")

# Plot for Percentages
plot_percentages <- plot_ly(uhf42_percentages, 
                            x = ~geography, y = ~Percent, 
                            color = ~Commute_Mode, type = "bar",
                            colors = custom_colors) %>%
  layout(title = list(text = "Commuting Mode Percentages by UHF 42 Neighborhoods", font = list(size = 9)),
         xaxis = list(title = "Neighborhood", tickangle = 45),
         yaxis = list(title = "Percentage", ticksuffix = "%", range = c(0, 100)),
         barmode = "stack",
         legend = list(title = list(text = "Commute Mode"),
                       font = list(size = 7)),
         font = list(size = 7))
plot_percentages         
```


