---
title: "final data"
output: html_document
---

```{r}
library(tidyverse)
library(janitor)
library(knitr)
library(gridExtra)
library(reshape2)
library(plotly)
```

```{r}
#import and clean data
overweight2022_df = 
  read.csv("data/overweight2022.csv", na = c("NA",".","")) |> 
   janitor::clean_names()

overweight_df <- read_csv("data/overweight2022.csv") %>%
  clean_names() %>%
  mutate(
    obesity_rate_raw = percent,  
    percent_numeric = sub("\\s*\\(.*$", "", percent),  
    obesity_rate = as.numeric(percent_numeric),  
    number_clean = as.numeric(gsub(",", "", number))  
  ) %>%
  filter(geo_type_desc == "Borough") %>%
  select(-percent_numeric)  
na_check <- overweight_df %>% filter(is.na(obesity_rate))
print(na_check)

overweight_overall_df <- read_csv("data/overweightoverall.csv") %>%
  clean_names() %>%
  mutate(
    obesity_rate = as.numeric(sub("\\s*\\(.*$", "", percent)),  
    geo_type = tolower(geo_type)  
  ) %>%
  filter(geo_type == "borough")  

```

```{r}
overweight2022_df <- overweight2022_df %>%
  mutate(obesity_rate = as.numeric(sub("\\s*\\(.*\\)$", "", percent))) 
borough_data <- overweight2022_df %>%
  filter(geo_type_desc == "Borough")

# Find the borough with the highest obesity rate
borough_with_highest_obesity <- borough_data %>%
  filter(obesity_rate == max(obesity_rate))
print(borough_with_highest_obesity)
```
Bronx has the highest overweight rate in 2022.


```{r}
#interactive plot for Obesity Rates and Numbers by Borough in NYC, 2022
library(ggplot2)
library(plotly)

obesity_plot <- ggplot(overweight_df, aes(x = geography, y = obesity_rate, fill = geography)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = number_clean),  
            vjust = -0.5, size = 4) +  
  labs(title = "Obesity Rates and Numbers by Borough in NYC, 2022",
       x = "Borough",
       y = "Obesity Rate (%)",
       fill = "Borough") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

interactive_plot <- ggplotly(obesity_plot)
interactive_plot

```



```{r}
# Create a line plot to show the trend of obesity rates by year for each borough
obesity_trend_plot <- ggplot(overweight_overall_df, aes(
  x = time_period,
  y = obesity_rate,
  group = geography,
  color = geography
)) +
  geom_line() +
  geom_point(aes(
    text = paste(
      "Borough:", geography,
      "<br>GeoID:", geo_id,
      "<br>Year:", time_period,
      "<br>Rate:", obesity_rate
    )
  )) +  
  labs(
    title = "Trend of Obesity Rates by Year and Borough",
    x = "Year",
    y = "Obesity Rate (%)",
    color = "Borough"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Convert ggplot to interactive plotly plot and enable custom tooltip
interactive_trend_plot <- ggplotly(obesity_trend_plot, tooltip = "text")

# Print the interactive plot
interactive_trend_plot

```

import commuting dataset
```{r}
commuting_df = 
  read.csv("data/commuting.csv", na = c("NA",".","")) |> 
   janitor::clean_names()
```

transfer datatype to numeric, filter by borough, using total number for each commuting mode
```{r}
borough_df = 
  commuting_df  |> 
  filter(geo_type_desc == "Borough") |> 
  mutate(
    across(c(bicycle_number, car_truck_or_van_number, public_transportation_number, walked_number),
                ~as.numeric(gsub(",", "", .))))
```

using long table to list the commuting mode for each borough
```{r}
commute_long <- pivot_longer(borough_df, 
                             cols = c(bicycle_number, car_truck_or_van_number, public_transportation_number, walked_number), 
                             names_to = "Method", 
                             values_to = "Count",
                             names_pattern = "(.*)_number")  
```

data explore for some graphs

```{r}
p1_plotly <- plot_ly(data = commute_long, x = ~geography, y = ~Count, type = 'bar', color = ~Method, colors = RColorBrewer::brewer.pal(8, "Set2"),
                     hoverinfo = 'text', text = ~paste("Borough: ", geography, "<br>Geo ID: ", geo_id, "<br>Method: ", Method, "<br>Count: ", Count),
                     name = 'Count of Commuters') %>%
  layout(title = 'Commuting Methods by Borough',
         xaxis = list(title = 'Borough'),
         yaxis = list(title = 'Count of Commuters'),
         barmode = 'group')

p1_plotly 
```

do the same for commuting modes percent
```{r}
commute_percent_long <- pivot_longer(borough_df, 
                                     cols = c(bicycle_percent, car_truck_or_van_percent, public_transportation_percent, walked_percent), 
                                     names_to = "Method", 
                                     values_to = "Percent",
                                     names_pattern = "(.*)_percent")
```


```{r}
p2_plotly <- plot_ly(data = commute_percent_long, x = ~geography, y = ~Percent, type = 'bar', color = ~Method, colors = RColorBrewer::brewer.pal(8, "Set1"),
                     hoverinfo = 'text', text = ~paste("Borough: ", geography, "<br>Geo ID: ", geo_id, "<br>Method: ", Method, "<br>Percentage: ", sprintf("%.1f%%", Percent)),
                     name = 'Percentage of Commuters') %>%
  layout(title = 'Percentage Distribution of Commuting Methods by Borough',
         xaxis = list(title = 'Borough'),
         yaxis = list(title = 'Percentage (%)', tickformat = ",.0%"),
         barmode = 'stack')

p2_plotly
```

we can see from the bar chart, public transportation is the most popular commuting mode in most borough, except for the Staten Island, where the car/truck/van is the most popular commuting mode.

```{r}
commuting_data = borough_df  |> 
  pivot_longer(
    cols = -c(time_period, 
              geo_type_desc, 
              geo_id, geo_rank, geography), 
    names_to = c("method", ".value"),
    names_pattern = "(.*)_(number|percent)$"
  )
```


